---
title: "assignment_6_ONEIL"
format: gfm
---

# Assignment 6: Data import and tidy data

```{r}
#| message: false
library(tidyverse)
library(knitr)
library(lubridate)
```

## Exercise 1. Tibble and Data Import

Import the data frames listed below into R and [parse](https://r4ds.had.co.nz/data-import.html#parsing-a-vector) the columns appropriately when needed. Watch out for the formatting oddities of each dataset. Print the results directly, **without** using `kable()`.

**You only need to finish any three out of the five questions in this exercise in order to get credit.**

#### 1.1 Create the following tibble manually, first using `tribble()` and then using `tibble()`. Print both results. \[We didn't have time to cover this in class, but look up how these functions work [here](https://r4ds.had.co.nz/tibbles.html#creating-tibbles)\]

`tibble():`

```{r}
tibble(a = c(1,2),b= c(2.1, 3.2),c= c("apple", "orange")) |> 
  print()
  
```

`tribble():`

```{r}
tribble(
  ~a, ~b, ~c,
  1, 2.1, "apple",
  2, 3.2, "orange"
) |> 
  print()
```

#### 1.2 Import `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset2.txt` into R. Change the column names into “Name”, “Weight”, “Price”.

```{r}
fruit<- read.csv("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset2.txt", header = FALSE)

colnames(fruit)<- c("Name", "Weight", "Price")

tibble(fruit)
```

#### 1.3 Import `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset3.txt` into R. Watch out for the first few lines, missing values, separators, quotation marks, and deliminaters.

```{r}
#| message: false 
fruit_1.3 <-read_delim("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/dataset3.txt", quote ="/", skip = 2, na = c("Not Available", "?"))
 tibble(fruit_1.3)
```

## Exercise 2. Weather station

This dataset contains the weather and air quality data collected by a weather station in Taiwan. It was obtained from the Environmental Protection Administration, Executive Yuan, R.O.C. (Taiwan).

#### 2.1 Variable descriptions

-   The text file `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/2015y_Weather_Station_notes.txt` contains descriptions of different variables collected by the station.

-   Import it into R and print it in a table as shown below with `kable()`.

```{r}
weather_station <- read.delim("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/2015y_Weather_Station_notes.txt", header = TRUE, sep = "-", dec = ".")

kable(weather_station)


```

`#` indicates invalid value by equipment inspection\
`*` indicates invalid value by program inspection\
`x` indicates invalid value by human inspection\
`NR` indicates no rainfall\
blank indicates no data

#### 2.2 Data tidying

-   Import `https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/2015y_Weather_Station.csv` into R. As you can see, this dataset is a classic example of untidy data: values of a variable (i.e. hour of the day) are stored as column names; variable names are stored in the `item` column.

-   Clean this dataset up and restructure it into a tidy format.

-   Parse the `date` variable into date format and parse `hour` into time.

-   Turn all invalid values into `NA` and turn `NR` in rainfall into `0`.

-   Parse all values into numbers.

-   Show the first 6 rows and 10 columns of this cleaned dataset, as shown below, *without* using `kable()`.

*Hints: you don’t have to perform these tasks in the given order; also, warning messages are not necessarily signs of trouble.*

```{r}
weather_2.2 <- read.csv("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/2015y_Weather_Station.csv")
```

Before cleaning:

```{r}

tibble(weather_2.2[1:6, 1:10])

```

After cleaning:

```{r}
#| warning: false
weather_raw <- read_csv("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/master/datasets/2015y_Weather_Station.csv", col_names = (TRUE), guess_max = Inf)


weather_clean <- weather_raw |>
  pivot_longer(cols = c("00":"23"), names_to = "hour", values_to = "value") |>
  pivot_wider(names_from = item, values_from = value) |>
  mutate(date = as.Date(date), across(AMB_TEMP:PM10, as.numeric), hour = hms::as_hms(paste0(hour, ":00:00")))


weather_clean <- weather_clean |>
  mutate(RAINFALL = ifelse(RAINFALL == "NR", 0, RAINFALL)) |>
  mutate(across(AMB_TEMP:PM10, ~na_if(.x, -99.99))) |>
  mutate(across(AMB_TEMP:PM10, as.numeric))

tibble(weather_clean[1:6, 1:10])

```

#### 2.3 Using this cleaned dataset, plot the daily variation in ambient temperature on September 25, 2015, as shown below.

```{r}
weather_clean |>
  filter(date == "2015/09/15") |>
  ggplot()+
  geom_line(aes(x = hour, y = AMB_TEMP, group = station))
```

#### 2.4 Plot the daily average ambient temperature throughout the year with a **continuous line**, as shown below.

```{r}
weather_clean |> 
  group_by(date) |> 
  summarise(daily_avg_temp = mean(AMB_TEMP, na.rm = TRUE)) |> 
  ggplot(aes(x = date, y = daily_avg_temp)) +
  geom_line()
```

#### 2.5 Plot the total rainfall per month in a bar chart, as shown below.

*Hint: separating date into three columns might be helpful.*

```{r}
#| warning: false
weather_clean <- weather_clean |>
  mutate(month = month(date), year = year(date))


weather_clean |> 
  group_by(month) |> 
  mutate(RAINFALL = as.numeric(RAINFALL)) |> 
  summarise(total_rainfall = sum(RAINFALL, na.rm = TRUE)) |> 
  ggplot(aes(x = month, y = total_rainfall)) +
  geom_bar(stat = "identity", fill = "black")
```

#### 2.6 Plot the per hour variation in PM2.5 in the first week of September with a **continuous line**, as shown below.

*Hint: uniting the date and hour and parsing the new variable might be helpful*

```{r}
#| warning: false
weather_clean$month <- as.character(weather_clean$month)
weather_clean |> 
  mutate(PM2.5 = as.numeric(PM2.5)) |> 
  mutate(day = day(date), month = month(date), year = year(date)) |> 
  mutate(datetime = as.POSIXct(paste(date, hour), format="%Y-%m-%d %H:%M:%S")) |> 
  filter(month == 9) |> 
  filter(between(day, 1, 7)) |> 
  ggplot()+
  geom_line(aes(x = datetime, y = PM2.5))
```
